\sys: Property-Guided Program Regeneration\\{\vspace{1em}\large Technical University of Denmark, MSc Thesis {\\\vspace{-1em}} Advisors: Christian Gram Kahlauge and Nikos Vasilakis

### 2025-07-31 docs/paper/draft.tex
Fix typos
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

\heading{A compromised javascript library}
The \ttt{flatmap-stream} library was implicated in a high-profile
supply-chain attack that exfiltrated Bitcoin credentials~\cite{ev:eurosec:2022}.
While the library preserved
its client-observable behavior, an malicious payload inserted in the module's code by an adversarial developer accessed the filesystem, network, and
global state under specific conditions.
Speficially, 
the library would
(1) detect whether it is running under the dependency tree of a popular bitcoin wallet
application~\cite{copay} and on the live bitcoin network,
(2) collect the user's private keys, 
(3) send those to an attacker-controlled machine,
and finally (4) perform its benign functionality.
Program regeneration can automatically remove such malicious payloads
using the insight that the insight that they retain the original library's functionality,
while performing a side-effectful attack on-the-side~\cite{harp:ccs:2021}.
This approach can secure the component even when the attack is highly obfuscated and 
activates under precise conditions.
The slightly simplified source code of the compromised library showing the malicious payload and a final
line that performs the benign flatmap functionality is shown below:
\begin{minted}[frame=lines]{js}
module.exports = function (stream, fn) {
 if (process.env.PRODUCTION) {
  const priv = copay.getKeys()?.[0]?.priv;
  if (priv && wallet.balance > 0) {
    http.request({ hostname: 'evil.com' })
        .send({ priv });
  }} return flatten(steam).apply(fn); };
\end{minted}

\begin{wrapfigure}[3]{r}{.7\columnwidth}
\vspace{-10pt}
\hspace{-10pt}
\begin{minted}{prolog}
graph(p_, I, O) :- graph(p, I, O).
fsignt(p_, S) :- fsignt(p, S).
pure(p_).
\end{minted}
\end{wrapfigure}
To the right is a program written in the \sys query language to express 
this regeneration goal.
This query requires the regenerated program \ttt{p\_} to exhibit the same I/O behavior as the original program \ttt{p}
but forbids side-effects.
\sys takes this query and arrives at the set of properties to extract from the
original program, given its knowledge base and the query.
This set is $\{$\ttt{graph(p, I, O)}, \ttt{inpt(P, I)}, \ttt{fsignt(P, S)}, \ttt{pure(P)}$\}$, where
\ttt{input(p, I)} means that \sys will use the \ttt{input} plugin extract a set of inputs from the original program
by providing the program's source code to an LLM instance and asking it to generate a set of inputs;
\ttt{graph(p, I, O)} extracts the client-observable I/O behavior of the program
by executing the original program on the generated inputs and recording the outputs
(an example I/O pair is $\langle\ttt{([1,[2,3]], (x)=>x+1)}\to\ttt{[2,3,4]}\rangle$);
and \ttt{fsignt(p, S)} extracts the function's signature using a language-aware parser.
The, \ttt{pure(p)} property means that \sys will not provide to the LLM 
instance any information that could encode side-effects, such as the
program's source code, a system call trace, or a filesystem access trace.
After 30.4 seconds and one attempt, having generated 30 I/O pairs, \sys generates the following program:
\begin{minted}[frame=lines]{js}
module.exports = function(stream, fn) {
  return stream.map(fn); };
\end{minted}

@@

\heading{Idea of program projection}
\sys can be viewed as a projection system, as shown in \cref{fig:projection}.
\sys projects a program into an abstract space of properties, 
then, it performs transformations within that space (the input query),
and finally, it regenerates this projection back into a concrete program.
The idea of casting programs into a high-dimensioanl space has been studied before~\cite{ben-nun2018neural,alon2019code2vec,ellis2020dreamcoder,nye2021platypus},
but \sys's perspective might offer interesting insights.

@@

\begin{wrapfigure}[5]{r}{0.4\columnwidth}
\begin{minted}{prolog}
graph(p_, I, O) :-
graph(p, I, O).
\end{minted}
\end{wrapfigure}
The query for \sys is shown on the right and asserts the the input/output behavior of the
original program is preserved in the output.
The resulting program must match the input-output behavior of the original.
The top listing shows the original \ttt{leftPad} library, and the second listing shows the output
of \sys.
Regeneration here matches the client-visible behavior of the original but misses
implementation details, such as the use of a cache for common
use-cases and the logarithmic complexity of the padding operation.
This example inspires future work on performance-aware regeneration.

@@

\heading{Fast Inverse Square Root}
The goal of this regeneration task is to transform an input 
program into a more idiomatic version that retains the same functionality.
The query for \sys is shown on the right and it uses the length of the output program as a proxy towards 
more idiomatic code.
\begin{wrapfigure}[4]{r}{0.7\columnwidth}
\begin{minted}{prolog}
graph(p_, I, O) :- graph(p, I, O).
#min N : len(p_, N)
\end{minted}
\end{wrapfigure}
The fast inverse square root function is a well-known algorithm that was
popularized by its use in the Quake III Arena source code.
It computes the inverse square root of a number using 
Newton's method through bit-level manipulation on the input floating-point number.
The original implementation is shown in \cref{lst:fast_inv_sqrt_original}.
\Cref{lst:fast_inv_sqrt_gpt} shows the output of naively regenerating the function using \gptmodel,
which retains the original's un-idiomatic use of magic numbers and bit-level manipulation.
\Cref{lst:fast_inv_sqrt_sys} has the output of \sys, which instead uses the standard library's
\ttt{sqrt} function to compute the inverse square root.
\sys provides only I/O examples to its backend LLM, and as such does not overfit to the original implementation.

>>>>

\heading{A compromised JavaScript library}
The \ttt{flatmap-stream} library was implicated in a high-profile
supply-chain attack that exfiltrated Bitcoin credentials~\cite{ev:eurosec:2022}.
While the library preserved
its client-observable behavior, an malicious payload inserted in the module's code by an adversarial developer accessed the filesystem, network, and
global state under specific conditions.
Specifically, 
the library would
(1) detect whether it is running under the dependency tree of a popular bitcoin wallet
application~\cite{copay} and on the live bitcoin network,
(2) collect the user's private keys, 
(3) send those to an attacker-controlled machine,
and finally (4) perform its benign functionality.
Program regeneration can automatically remove such malicious payloads
using the insight that the insight that they retain the original library's functionality,
while performing a side-effectful attack on-the-side~\cite{harp:ccs:2021}.
This approach can secure the component even when the attack is highly obfuscated and 
activates under precise conditions.
The slightly simplified source code of the compromised library showing the malicious payload and a final
line that performs the benign \ttt{flatmap} functionality is shown below:
\begin{minted}[frame=lines]{js}
module.exports = function (stream, fn) {
 if (process.env.PRODUCTION) {
  const priv = copay.getKeys()?.[0]?.priv;
  if (priv && wallet.balance > 0) {
    http.request({ hostname: 'evil.com' })
        .send({ priv });
  }} return flatten(steam).apply(fn); };
\end{minted}

\begin{wrapfigure}[3]{r}{.7\columnwidth}
\vspace{-10pt}
\hspace{-10pt}
\begin{minted}{prolog}
graph(p_, I, O) :- graph(p, I, O).
fsignt(p_, S) :- fsignt(p, S).
pure(p_).
\end{minted}
\end{wrapfigure}
To the right is a program written in the \sys query language to express 
this regeneration goal.
This query requires the regenerated program \ttt{p\_} to exhibit the same I/O behavior as the original program \ttt{p}
but forbids side-effects.
\sys takes this query and arrives at the set of properties to extract from the
original program, given its knowledge base and the query.
This set is $\{$\ttt{graph(p, I, O)}, \ttt{fsignt(P, S)}, \ttt{pure(P)}$\}$, where
\ttt{input(p, I)} means that \sys will use the \ttt{input} plugin extract a set of inputs from the original program
by providing the program's source code to an LLM instance and asking it to generate a set of inputs;
\ttt{graph(p, I, O)} extracts the client-observable I/O behavior of the program
by executing the original program on the generated inputs and recording the outputs
(an example I/O pair is $\langle\ttt{([1,[2,3]], (x)=>x+1)}\to\ttt{[2,3,4]}\rangle$);
and \ttt{fsignt(p, S)} extracts the function's signature using a language-aware parser.
The, \ttt{pure(p)} property means that \sys will not provide to the LLM 
instance any information that could encode side-effects, such as the
program's source code, a system call trace, or a filesystem access trace.
After 30.4 seconds and one attempt, having generated 30 I/O pairs, \sys generates the following program:
\begin{minted}[frame=lines]{js}
module.exports = function(stream, fn) {
  return stream.map(fn); };
\end{minted}

@@

\heading{Idea of program projection}
\sys can be viewed as a projection system, as shown in \cref{fig:projection}.
\sys projects a program into an abstract space of properties, 
then, it performs transformations within that space (the input query),
and finally, it regenerates this projection back into a concrete program.
The idea of casting programs into a high-dimensional space has been studied before~\cite{ben-nun2018neural,alon2019code2vec,ellis2020dreamcoder,nye2021platypus},
but \sys's perspective might offer interesting insights.

@@

\begin{wrapfigure}[5]{r}{0.4\columnwidth}
\begin{minted}{prolog}
graph(p_, I, O) :-
graph(p, I, O).
\end{minted}
\end{wrapfigure}
The query for \sys is shown on the right and asserts the input/output behavior of the
original program is preserved in the output.
The resulting program must match the input-output behavior of the original.
The top listing shows the original \ttt{leftPad} library, and the second listing shows the output
of \sys.
Regeneration here matches the client-visible behavior of the original but misses
implementation details, such as the use of a cache for common
use-cases and the logarithmic complexity of the padding operation.
This example inspires future work on performance-aware regeneration.

@@

\heading{Fast Inverse Square Root}
The goal of this regeneration task is to transform an input 
program into a more idiomatic version that retains the same functionality.
The query for \sys is shown on the right and it uses the length of the output program as a proxy towards 
more idiomatic code.
\begin{wrapfigure}[4]{r}{0.7\columnwidth}
\begin{minted}{prolog}
graph(p_, I, O) :- graph(p, I, O).
#min N : len(p_, N)
\end{minted}
\end{wrapfigure}
The fast inverse square root function is a well-known algorithm that was
popularized by its use in the Quake III Arena source code.
It computes the inverse square root of a number using 
Newton's method through bit-level manipulation on the input floating-point number.
The original implementation is shown in \cref{lst:fast_inv_sqrt_original}.
\Cref{lst:fast_inv_sqrt_gpt} shows the output of naively regenerating the function using \gptmodel,
which retains the original's unidiomatic use of magic numbers and bit-level manipulation.
\Cref{lst:fast_inv_sqrt_sys} has the output of \sys, which instead uses the standard library's
\ttt{sqrt} function to compute the inverse square root.
\sys provides only I/O examples to its backend LLM, and as such does not overfit to the original implementation.
